---
title: "Visualization Lab 2"
author: "Roshni Sundaramurthy (rossu809); Brian Masinde (brima748)"
date: "17 September 2018"
output:
  html_document:
    df_print: paged
  html_notebook:
    theme: journal
  pdf_document: default
fontsize: 11pt
bibliography: references.bib
---

### Assignment 1. Perception Visualization

```{r data, echo = FALSE}
olive <- read.csv("olive.csv", sep = ",", header = TRUE)
```

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(plotly)
library(MASS)
library(gridExtra)
```

#### Question 1.
```{r question1_a, echo=FALSE}
p_vs_o <- ggplot(olive, aes(x=oleic, y=palmitic,
                     color = linolenic)) +
  geom_point() +
  ggtitle("Dependence of Palmitic on Oleic colored by numeric Linolenic") +
  theme_minimal()
  
p_vs_o
```

```{r cut_interval}
olive$linolenic2 <- cut_interval(olive$linolenic, n=4)
```

The plots demonstrate the preattentive perception problem. With the first plot it is not easy to make out boundaries while in the second plot one can easily perceive boundaries.

```{r question1_b, echo=FALSE}
p_vs_o2 <- ggplot(olive, aes(x=oleic, y=palmitic, color = linolenic2)) + 
  geom_point() +
  ggtitle("Dependence of Palmitic on Oleic colored by discretized Linolenic") +
  theme_minimal()
  #scale_fill_manual(name = "Discrete Linolenic")

p_vs_o2
```

#### Question 2.
It is more difficult to differentiate between the categories using angle compared to using color and size. The spoke diagram requires a combination of line length (2.8 bits) and line orientation (3 bits) which is more than size (2.2 bits) and color brightness (2.1 bits)

```{r question 2_a, warning=FALSE, echo=FALSE}
pal_vs_ole_col <- ggplot(olive, aes(oleic, palmitic)) +
  geom_point(aes(color=linolenic2))+
  ggtitle("Oleic vs Palmitic colored by Discretized Linolenic") +
  theme_minimal()

pal_vs_ole_col
```


```{r question 2_b, warning = FALSE, echo=FALSE}
pal_vs_ole_size <- ggplot(olive, aes(oleic, palmitic)) +
  geom_point(aes(size=linolenic2)) +
  ggtitle("Oleic vs Palmitic sized by Discretized Linolenic") +
  theme_minimal()

pal_vs_ole_size
```

```{r question 2_c, warning=FALSE}
pal_vs_ole_angle <- ggplot(olive, aes(oleic, palmitic)) +
  geom_point() +
  geom_spoke(aes(angle=as.numeric(linolenic2),radius= 35)) +
  ggtitle("Palmitic vs Oleic angle by Linolenic") +
  theme_minimal()

pal_vs_ole_angle
```

#### Question 3. 
For the first plot it is difficult to judge boundaries because the color takes a continuous scale, it requires more attention to differentiate the color scales (attentive). For the second graph with region as factors it is quick to judge boundaries hence preattentive.

```{r Question3, echo=FALSE}
o_vs_e <- ggplot(olive, aes(x= eicosenoic, y = oleic, color = Region)) + 
  geom_point() +
  ggtitle(" Oleic vs Eicosenoic colored by Region") +
  theme_minimal()
o_vs_e
```

```{r question3_b, echo=FALSE}
o_vs_e0 <- ggplot(olive, aes(x=eicosenoic, y = oleic , color = as.factor(Region)))+
  geom_point() +
  ggtitle("Oleic vs Eicosenoic colored by factored Region") +
  theme_minimal()

o_vs_e0
```

#### Question 4

This is an attentive perception problem. It is quite difficult to process information and judge the boundaries on the three groups. One has to analyse color shape and size and position on the X and Y axis at the same time.
 
```{r discretization, echo=FALSE}
olive$linoleic2 <- cut_interval(olive$linoleic, n=3)
olive$palmitic2 <- cut_interval(olive$palmitic, n=3)
olive$palmitoleic2 <- cut_interval(olive$palmitoleic, n=3)
```

```{r question4, warning=FALSE}
o_vs_e1 <- ggplot(olive, aes(x=eicosenoic, y= oleic)) + 
  geom_point(aes(color=linoleic2, size=palmitoleic2, shape=palmitic2)) +
  ggtitle("Oleic vs Eicosenoic")+
  theme_minimal()

o_vs_e1
```

#### Question 5.
The color dominates every other aesthetic hence it makes it easier to notice the boundaries despite shape and size. Treisman's theory [@treisman] on visual analysis "search for the presence of a visual primitive is automatic and
parallel, whereas search for the absence of the same feature is serial and requires focused
attention". 

```{r question5, warning=FALSE}
o_vs_e2 <- ggplot(olive, aes(x=oleic, y=eicosenoic)) + 
  geom_point(aes(color=Region, size=palmitoleic2, shape=palmitic2))

o_vs_e2
```

#### Question 6.

Without labels on the individual sections, the different areas appear to have the same proportions. This demonstrates the relative judgement problem.
```{r question6, echo=FALSE}
plot_ly(data = olive, labels=~Area, type = 'pie', showlegend = TRUE, textinfo = "text", text = "") %>%
  layout(title = 'Proportion of Oils from different regions',
        xaxis = list(showgrid = F, zeroline = F, showticklables = F),
         yaxis = list(showgrid = F, zeroline = F, showticklables = F))

```


#### Question 7.

The contour plot shows that there are observations in areas where in actual scatter plot there aren't any. The contour plot shows that there are five distinct density areas (two sparsely dense and three highly dense).

```{r question7}
l_vs_e <- ggplot(olive, aes(x = eicosenoic, y = linoleic)) +
  geom_density_2d()+
  theme_minimal() +
  ggtitle("Contour plot of linoleic vs eicosenoic")
  
l_vs_e  
```


```{r question7_b, echo=FALSE}
l_vs_e1 <- ggplot(olive, aes(x=eicosenoic, y = linoleic))+
  geom_point() +
  ggtitle("linoleic vs Eicosenoic Scatter plot ") +
  theme_minimal()

l_vs_e1
```


### Assignment 2: Multidimensional Scaling of a high-dimensional object

```{r base_data, echo= FALSE}
# reading the data
baseball <- read.csv("baseball-2016.csv", sep = ",", header = TRUE)
```

#### Question1
The xlsx file is converted to a csv file, makes it easier to load the data without extra packages. It is reasonable to scale the data because they have different ranges so that features with large scales do not dominate. Example, comparing maximum value and minimum value for "AB" and "Home runs per game". The two are on totally different scales.

```{r range_ab}
# range of AB
range1 <- max(baseball$AB) -  min(baseball$AB)
range1
```

```{r range_hrpg}
# range of HR per game
range2 <- max(baseball$HR.per.game) - min(baseball$HR.per.game)

range2
```

The density plot below graphically explains why scaling is needed. Appropriate density should be similar to log normal density.

```{r base_density}
plot(density(as.matrix(baseball[3:ncol(baseball)])), main = "Density of baseball numeric variables ")
```


#### Question 2

```{r scale, echo=FALSE}
# scale the 26 numeric columns
baseball2 <- scale(baseball[ ,3:ncol(baseball)])
```

```{r mink}
# Getting distance between points using Minkowski
mink_d <- dist(baseball2, method = "minkowski", p =2)
```

```{r isoMDS, message=FALSE, warning=FALSE, paged.print=FALSE}
# get 2 column vector with fitted configuration 
resid <- isoMDS(mink_d, k=2 )

# getting the coordinates
coords <- resid$points

# convert coords to dataframe
coordsMDS <- as.data.frame(coords)

# cbind to get column Team and League
coordsMDS2 <- cbind(baseball[ , 1:2], coordsMDS)
```

```{r var, echo = FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# plot the new dimension and color by  League
plot_ly(coordsMDS2, x= ~V1, y = ~V2, type ="scatter", hovertext = ~ Team,
            color = ~League, colors = c("#0444BF","#F46A4E"))
```

There appears to be a difference between the two leagues; most of the NL league teams are on the second and third quadrant (negative V2) while most of the AL league teams are on the first and fourth quadrant (positive V2). The y-axis "V2" has the best differentiation between the leagues. The Boston Red Sox, Atlanta Braves and Philadelphia Philies appear to be outliers. The AL teams are closely clustered together compared to the NL teams.

```{r question3_2}
# Shepard diagrams assess the goodness-of-fit of MDS techniques
# run shep on points from isoMDS and using same distance as calculated in previous step
shep <- Shepard(mink_d, coords)
```

```{r num_mink, echo=FALSE}
# convert the distance to numeric
delta <- as.numeric(mink_d)
```


```{r num_d_coords, echo=FALSE}
# All the coords in one column
D <- as.numeric(dist(coords))
```

```{r index1, echo=FALSE}
# create a square matrix with dimensions from the rows of coords
# n as rows 
n = nrow(coords)

# index as empyt matrix
index <- matrix(1:n, nrow = n, ncol = n)

# get index of the lower triangle of the index matrix and convert it to numeric
index1 <- as.numeric(index[lower.tri(index)])

```

```{r index2, echo=FALSE}
# same procedur
n = nrow(coords)

index <- matrix(1:n, nrow=n, ncol=n, byrow = T)

index2 <- as.numeric(index[lower.tri(index)])
```

```{r rownames, echo=FALSE}
# i dont want to edit original data
baseball3 <- baseball

rownames(baseball3) <-baseball3[ ,1]

```

```{r plot_s}
# plotly
plot_ly()%>%
  add_markers(x=~delta, y=~D, hoverinfo = 'text',
        text = ~paste('Obj1: ', rownames(baseball3)[index1],
                      '<br> Obj2: ', rownames(baseball3)[index2])) %>%
  #if nonmetric MDS inolved
  add_lines(x=~shep$x, y=~shep$yf) %>%
  layout(title = "Shepard's plot")
```


```{r question4_2, echo=FALSE}
# get MDS variable V2 and cbind with points from isoMDS
d4 <- cbind(baseball[ ,1:2], coords[,2], baseball2[,])

# rename col3 to V2
colnames(d4)[3] <- c("V2")
```


```{r fxn2, echo=FALSE}
# function to successively get plots of V2 vs other variables
myplot <-function(df, y_string){
  ggplot(df, aes_string(x = "V2", y = y_string, color = "League"))+
    geom_point()+
    geom_vline(xintercept = 0, color = "black")+
    geom_hline(yintercept = 0, color = "black")
}
```

```{r hrpg}
# Variables that seem to have strongest postive relationship
# HR per game:
myplot(d4, "HR.per.game")
```

```{r x3b}
# X3B
myplot(d4, "X3B")
```

Per the plot, HR per game and HR have the same scatter points. The two have strongest positive connection to V2 while X3B has the strongest negative connection to V2.

### Appendix

```{r ref.label=knitr::all_labels(), echo = T, eval = F}

```

### References

